### 2011-05-31 thesis/contents.tex
A lot of things on how to cope with multiple novel variables.

\section{Detecting New Classes}
This section focus on detecting a novel class on single variable of the
conceptual map. As seen before on \autoref{chap:novelty-intro},
novelty detection can be done by implementing a threshold on
$P(x|\overline{novel})/P(x)$.
In order to correctly measure novelty on a single variable both
conditional and unconditional probabilities need to be coherently defined:

The conceptual layer build a \emph{probabilistic graphical model} $G$ that
represents the distribution of the sensed features assuming that both the known
structure and the known classes hold in the sample.
For that reason the correct approach is to use $P_G(x)$ as an approximation for
$P(x|\overline{novel})$.

Assume now that everything the conceptual layer knows holds true except the
class of a single variable $a$, which the system possess no information on. 
In that case the correct approach is to replace the factors associated with
that variable with factors representing the prior-information the system has on
the unconditional distribution of those factors.
In case no information is available this means replace by a uniform factor\footnote{In case the
unconditional factors are uniform, this is the same as omitting the factors
from the graph structure.} accordingly to the principle of maximum entropy.
Note that in case some information was known on the prior distribution, the
correct approach is to replace the factors by the distributions that have that
information in account (as is the case in \autoref{sec:unlabelled-data}).
Let this new factor graph with unconditional factors $\phi_u$ on the variable
$a$ be $U$. $P(x)$ is then approximated by $P_U(x)$.

@@

\section{Copping with Multiple Novel Variables}
On the previous section it was shown how to detect a novel class on a variable
assuming that the rest of the graph variables and structure is correctly modeled
by the agent knowledge.
But the test to detect a given variable $a$ is novel requires both a structure
$G$ representing the case where no knowledge gaps exist and a structure $U$
representing the case where $a$ was identified as a novel class.

It may be tempting to use a deterministic approach on the system such as:
when the agent determines a variable to have a novel class, it marks it as novel
and proceeds.

@@

In this section we show how to handle novelty in several variables
in simultaneous.\footnote{I am almost sure the math here will work out the
way I think}

>>>>

\section{Detecting Novelty on a Single Variable}
This section focus on detecting a novel class on single variable of the
conceptual map. As seen before on \autoref{chap:novelty-intro},
novelty detection can be done by implementing a threshold on
$P(x|\overline{novel})/P(x)$.
In order to correctly measure novelty on a single variable both
conditional and unconditional probabilities need to be coherently defined:

The conceptual layer builds a \emph{probabilistic graphical model} $G$ that
represents the distribution of the sensed features $x$ assuming that both the
known structure and the known classes of all variables hold in the sample.
The correct approach is then to use $P_G(x)$ as an approximation for
$P(x|\overline{novel})$.

Assume now that everything the conceptual layer knows holds true except the
class of a single variable $a$, which the system possess no information on. 
In that case, the correct approach is to replace the factors associated with
$a$ with factors representing the prior-information the system has on
the unconditional distribution of those.
Let this new graph with unconditional factors on the variable
$a$ be $U_a$. $P(x)$ is then approximated by $P_{U_a}(x)$.

Note that in case no prior-information is available on the unconditional
factors this means replacing them by a uniform factor\footnote{In case the
unconditional factors are uniform, this is the same as omitting the factors
from the graph structure.}. In case some information is known
(i.e. unlabelled data) the correct approach is to incorporate it into the
unconditional factors using the principle of maximum entropy as seen on
\autoref{sec:unlabelled-data}.

@@

Having now defined $P_G(x)$ and $P_{U_a}$ as models models for
$P(x|\overline{novel})$ and $P(x)$ (\autoref{fig:nd-single-variable}),
$P_G(x)/P_{U_a}(x)$ is used to define an order-relation on the novelty of $a$.
Now all that is missing is a threshold $t_A$ to trigger on.
Such a threshold can be obtained by calculating which value of $t_A$
triggers a desired amount of the labelled data.
Also, since there is no reason to distinguish between different instances of
a variable type, only one threshold by variable type needs to be trained.

Concluding: novelty on a variable $a$ of type $A$ should be detected when:

\begin{equation}
\frac{P_G(x)}{P_{U a}(x)} < t_A
\end{equation}

where $t_A$ is a threshold for variables of type $A$, $P_G(x)$ denotes the
probability of the sensed variables $x$ being generated by a distribution
where $a$ is known, and $P_{Ua}$ denotes the probability of $x$ being generated
by a distribution where only $a$ is novel.

\section{Copping with Multiple Novel Variables}
On the previous section it was shown how to detect a novel class on a variable
assuming that the rest of the graph variables and structure is correctly modeled
by the agent knowledge.
In this section methods are presented that not only allow to extended this to
detection on multiple variables but also allow the system to incorporate
novelty information back in the graphical model in order to continue its
probabilistic and uncertain reasoning.

When trying to extend the method to deal with multiple novel variables it may
be tempting to follow a deterministic approach such as:
when the agent determines a variable $a$ to have a novel class, it marks it as
novel and proceeds from that moment by using a graphical model where the factors
associated with the $a$ replaced by unconditional factors.
But such an approach would not be able to optimally detect novel variables
as later decisions are dependent on former ones.
It would also not be able to adapt itself and reconsider previously decisions
when more data becomes available for the agent.

To handle those issues a fully probabilistic approach should be taken
instead of a deterministic.
That can be done by extending all the variables on the factor graph representing
the semantic map with an additional class \emph{novel} and extending all the
factors with information on the unconditional distribution.

Such a factor graph would be able to handle novel variables and incorporate
prior-information on the unconditional distributions since it is able to
represent any state of the $X$ variables present on the graph where any
subset of them is novel. But in order to add the ability to control the
novelty thresholds a normalization factor $\phi_N$ needs to be introduced.

In order to give full control on the threshold on which any subset of
variables is considered novel the factor $\phi_N$ needs to be dependent on all
the $X$ variables and act as a weighting function that correctly scales the
likelihood of a given instance of the $X$ variables in order to make a
\emph{MAP} operation on the graph to yield the desired value.
Such a factor would require tunning $2^N$ parameters, describing how to
distribute the weights over all the possible $2^N$ subsets of the $X$ variables.

Aiming at feasibility a simpler approach should be used.
Requiring

This is the factor $\phi_N(X) = \prod_{i \in X} t_i$.
As so the factor $\phi_N$ can be decomposed into several $\phi_{N_i} = t_i$, one
for each variable, that measure on the threshold of detecting novelty on
the single variable $i$.

More formally, there is an interest in the new graph
Nonetheless after that a mechanism to allow to control the thresholds on
detecting novelty also needs to simple doing that the ability to control the thresholds on
variables is lost
In order to measure novelty of a set of variable the full set of approach
Aiming at obtaining a correct behaviour for detecting novelty on a certain
subset of variables
Optimally the system should re-evaluate novelty for all subsets of variables
each time new information is added.
A method to define a threshold on any subset of the $N$ variables also needs to
be introduced.

After that the system would be still be able to reason over variables on the
graph and deduce extra new novel variables if it was the case.
But such an approach would be dependent on the order of variables that the
system considers for novelty detection. And it would also suffer from inability
to reconsider the novelty on any given variable as more data becomes available.

The used approach can be extended to multiple variables. Though in order to
expand it to all the $n$ variables in the graph, $n!$ unconditional
probabilities would need to be evaluated.

Aiming at a fully probabilistic integration of novelty detection on any
number of variables the following method was introduced:

But the test to detect a given variable $a$ is novel requires both a structure
$G$ representing the case where no knowledge gaps exist and a structure $U$
representing the case where $a$ was identified as a novel class.

@@
